---
output:
  pdf_document: default
  html_document: default
---
Announcements

1. Online Midterm Tuesday, Oct 19 covering material up through this week.
2. Revise and resubmit policy for project - due in a week.

123 GO - What's your beverage of the day?

Problem: If your code doesn't work, then it's hard to write a report.
Options:

1. Push through it - could recover points on previous assignments
2. Give you code that works - no way to recover missed points on previous assignments
3. ~~No report, just party~~
4. Group project - 2 people


## markdown demo

WYSIWYG vs. markup
"What you see is what you get"

markup languages - code to represent the document.

- HTML hypertext markup language
- markdown
- LaTeX
- XML extensible markup language
- Maybe any tech acronym that ends in ML?

$$
\LaTeX: \quad y = ax + b
$$

123 GO: What's an __advantage__ of a markup language?

- more control
- less control- true for markdown
- generating dynamic output

1. markdown is designed for simplicity.
2. I love markdown.
1. You should use it for your report!
4. Can't be bothered to count.

Try to knit an Rmd and see if it works.

What's the relationship between Rmd (Rmarkdown) and markdown?

markdown is general purpose
Rmarkdown is R specific.

`code` like that.

Chunk of code to run:

Inside the curly braces are chunk options.
```{r,  echo=FALSE}
n = 100
n
x = rnorm(n)
plot(x)
```


Put R code inline. The value of n is `r n`.

123 GO: What problem does inline code solve / prevent?

It helps to make computation reproducible.

# Big text

### Smaller, but still a header.

#### Smallest


Example of a link: [alt (display) text](URL).

Notes: [datalab](https://ucdavisdatalab.github.io/workshop_r_basics/exploring-your-data.html#data-visualization)

```{r}
head(iris)
```

Pretty it up:

```{r}
knitr::kable(head(iris))
```

Directly making a table:

col1 | col2 | col3
- | -- | ---
a | b | 1
c | d | 2




## ggplot2

Notes: [datalab](https://ucdavisdatalab.github.io/workshop_r_basics/exploring-your-data.html#data-visualization)

Three popular systems for creating visualizations in R:

The base R functions
The lattice package
The ggplot2 package

Pick one, don't mix and match.

Base R uses a "painter's canvas" model.


```R
n = 100
x = rnorm(n)
y = rnorm(n)
plot(x, y)

# 123 GO: Will the big yellow line cover the points?
abline(h = 0, lwd = 10, col = "yellow")
```

In contrast to base R, ggplot2 uses a "declarative" model based on the grammar of graphics.

The idea of a grammar of graphics is that visualizations can be built up in layers. In ggplot2, the three layers every plot must have are:

- Data
- Geometry
- Aesthetics


```R
earn = read.csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-23/earn.csv")
```

Let’s make a line plot that shows median earnings for each quarter in 2019, with separate lines for men and women.

```R
earn19 = earn[earn$year == 2019, ]
earn19 = earn19[earn19$race == "All Races" &
  earn19$ethnic_origin == "All Origins" &
    earn19$age == "16 years and over", ]
```



## Layer 1 - data

ggplot and most other Tidyverse packages are designed for working with tidy data frames. Tidy means:

1. Each observation has its own row.
2. Each feature has its own column.
3. Each value has its own cell.

```R
library(ggplot2)

ggplot(earn19)

```


## Layer 2 - geometry

```R
ggplot(earn19) + geom_line()
```


## Layer 3 - aesthetics

```R

ggplot(earn19, aes(x = quarter, y = median_weekly_earn, linetype = sex)) + 
geom_line()

```

If you want to set an aesthetic to a constant value, rather than one that’s data dependent, do so outside of the aesthetic layer. For instance, suppose we want to make the lines blue:

```R
ggplot(earn19, aes(x = quarter, y = median_weekly_earn, linetype = sex)) +
  geom_line(color = "blue")
```

## Layer 4 - scales

```R
ggplot(earn19, aes(x = quarter, y = median_weekly_earn, linetype = sex)) +
  geom_line() + labs(x = "Quarter", y = "Median Weekly Salary (USD)",
      title = "2019 Median Weekly Salaries, by Sex", linetype = "Sex")
```


## Saving

```R
ggsave("line.png")
```

The Plot Device
You can also save a plot with one of R’s “plot device” functions. The steps are:

1. Call a plot device function: png, jpeg, pdf, bmp, tiff, or svg.
1. Run your code to make the plot.
1. Call dev.off to indicate that you’re done plotting.

This strategy works with any of R’s graphics systems (not just ggplot2).

Here’s an example:

```R
# Run these lines in the console, not the notebook!
```


