---
output:
  pdf_document: default
  html_document: default
---
Announcements

1. Online Midterm Tuesday, Oct 19 covering material up through this week.
2. Revise and resubmit policy for project - due in a week.

123 GO - What's your beverage of the day?

Problem: If your code doesn't work, then it's hard to write a report.
Options:

1. Push through it - could recover points on previous assignments
2. Give you code that works - no way to recover missed points on previous assignments
3. ~~No report, just party~~
4. Group project - 2 people


## markdown demo

WYSIWYG vs. markup
"What you see is what you get"

markup languages - code to represent the document.

- HTML hypertext markup language
- markdown
- LaTeX
- XML extensible markup language
- Maybe any tech acronym that ends in ML?

$$
\LaTeX: \quad y = ax + b
$$

123 GO: What's an __advantage__ of a markup language?

- more control
- less control- true for markdown
- generating dynamic output

1. markdown is designed for simplicity.
2. I love markdown.
1. You should use it for your report!
4. Can't be bothered to count.

Try to knit an Rmd and see if it works.

What's the relationship between Rmd (Rmarkdown) and markdown?

markdown is general purpose
Rmarkdown is R specific.

`code` like that.

Chunk of code to run:

Inside the curly braces are chunk options.
```{r,  echo=FALSE}
n = 100
n
x = rnorm(n)
plot(x)
```


Put R code inline. The value of n is `r n`.

123 GO: What problem does inline code solve / prevent?

It helps to make computation reproducible.

# Big text

### Smaller, but still a header.

#### Smallest


Example of a link: [alt (display) text](URL).

Notes: [datalab](https://ucdavisdatalab.github.io/workshop_r_basics/exploring-your-data.html#data-visualization)

```{r}
head(iris)
```

Pretty it up:

```{r}
knitr::kable(head(iris))
```

Directly making a table:

col1 | col2 | col3
- | -- | ---
a | b | 1
c | d | 2





