
How was the DS panel?

Which data sounds most interesting to you?

1. COVID cases
2. public financial records
3. news [GDELT](https://www.gdeltproject.org/)
4. census data


## ggplot2

Notes: [datalab](https://ucdavisdatalab.github.io/workshop_r_basics/exploring-your-data.html#data-visualization)

Authoritative reference: [ggplot2 book](https://ggplot2-book.org/introduction.html)

> "ggplot2 provides beautiful, hassle-free plots that take care of fiddly details like drawing legends. 
> ...
> rather than spending your time making your graph look pretty, you can instead focus on creating the graph that best reveals the message in your data."


Three popular systems for creating visualizations in R:

The base R functions
The lattice package
The ggplot2 package

Pick one, don't mix and match.

Base R uses a "painter's canvas" model.


```{r}
n = 100
x = rnorm(n)
y = rnorm(n)
plot(x, y)

# 123 GO: Will the big yellow line cover the points?
abline(h = 0, lwd = 10, col = "yellow")
```

In contrast to base R, ggplot2 uses a "declarative" model based on the grammar of graphics.

The idea of a grammar of graphics is that visualizations can be built up in layers. In ggplot2, the three layers every plot must have are:

- Data
- Geometry
- Aesthetics


```{r}
earn = read.csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-23/earn.csv")
```

Let’s make a line plot that shows median earnings for each quarter in 2019, with separate lines for men and women.

```{r}
earn19 = earn[earn$year == 2019, ]
earn19 = earn19[earn19$race == "All Races" &
  earn19$ethnic_origin == "All Origins" &
    earn19$age == "16 years and over", ]

earn19 = earn[with(earn, year == 2019 &
                    race == "All Races" &
                    ethnic_origin == "All Origins" &
                    age == "16 years and over"
                ), ]

```

The tidyverse way:

```{r}
library(tidyverse)

earn19b = earn %>% filter(year == 2019,
                          race == "All Races",
                          ethnic_origin == "All Origins",
                          age == "16 years and over"
)
```




## Layer 1 - data

ggplot and most other Tidyverse packages are designed for working with tidy data frames. Tidy means:

1. Each observation has its own row.
2. Each feature has its own column.
3. Each value has its own cell.

If you've studied databases, this is 3rd normal form.

```{r}
library(ggplot2)

ggplot(earn19)

```

An example that came up for a student in the last project:

```{r}

library(tidyverse)

infections = read.csv("infections.csv", row.names = 1)

inf2 = pivot_longer(infections, everything(), names_to = "shape", values_to = "count_infected")

ggplot(inf2, aes(x = shape, y = count_infected)) + geom_boxplot()

```



## Layer 2 - geometry

```{r}
ggplot(earn19) + geom_line()
```


## Layer 3 - aesthetics

```{r}
ggplot(earn19, aes(x = quarter, y = median_weekly_earn, linetype = sex)) + 
geom_line()
```

If you want to set an aesthetic to a constant value, rather than one that’s data dependent, do so outside of the aesthetic layer. For instance, suppose we want to make the lines blue:

```{r}
ggplot(earn19, aes(x = quarter, y = median_weekly_earn, linetype = sex)) +
  geom_line(color = "blue")
```

## Layer 4 - scales

```{r}
p = ggplot(earn19, aes(x = quarter, y = median_weekly_earn, linetype = sex)) +
  geom_line() + 
  labs(x = "Quarter", y = "Median Weekly Salary (USD)",
      title = "2019 Median Weekly Salaries, by Sex", linetype = "Sex",
      caption = "Data from Nick Ulle's R Basics")
```

## Annotations

```{r}
p + annotate("text", x = 3, y = 1000, label = "elbow", vjust = "top")
```



## Saving

```{r}
ggsave("line.png")
```

The Plot Device
You can also save a plot with one of R’s “plot device” functions. The steps are:

1. Call a plot device function: png, jpeg, pdf, bmp, tiff, or svg.
1. Run your code to make the plot.
1. Call dev.off to indicate that you’re done plotting.

This strategy works with any of R’s graphics systems (not just ggplot2).

Here’s an example:

```{r}
# Run these lines in the console, not the notebook!
```


