
How was the DS panel?

Which data sounds most interesting to you?

picking something for class:

1. 1 COVID cases
2. 4 public financial records
3. 5 news [GDELT](https://www.gdeltproject.org/)
4. census data (2020 ACS)

Announcements:
Project is due tomorrow
Office hours tomorrow at 2.

## ggplot2

Notes: [datalab](https://ucdavisdatalab.github.io/workshop_r_basics/exploring-your-data.html#data-visualization)

Authoritative reference: [ggplot2 book](https://ggplot2-book.org/introduction.html)

> "ggplot2 provides beautiful, hassle-free plots that take care of fiddly details like drawing legends. 
> ...
> rather than spending your time making your graph look pretty, you can instead focus on creating the graph that best reveals the message in your data."


Three popular systems for creating visualizations in R:

The base R functions - we've used this.
The lattice package - skipping for now
The ggplot2 package - very popular, worth learning

base R uses a "painter's canvas" model.

123 GO: will the big red line "paint over" the points?
yes

```{r}
x = rnorm(100)
plot(x)
abline(h = 0, lwd = 10, col = "red")  # < a big red line

```



```{r}
earn = read.csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-23/earn.csv")
```

Let’s make a line plot that shows median earnings for each quarter in 2019, with separate lines for men and women.

Start with the right data.
```{r}
# 1
earn19 = earn[earn$year == 2019, ]
earn19 = earn19[earn19$race == "All Races", ]
earn19 = earn19[earn19$ethnic_origin == "All Origins", ]
earn19 = earn19[earn19$age == "16 years and over", ]
```

```{r}
# 2 with
earn19b = earn[with(earn, year == 2019 &
  race == "All Races" &
  ethnic_origin == "All Origins" &
  age == "16 years and over"), ]
```

The tidyverse way:
```{r}
# 3 dplyr filter
library(tidyverse)
earn19c = earn %>% filter(
  year == 2019, 
  race == "All Races",
  ethnic_origin == "All Origins",
  age == "16 years and over")
```

ggplot2

gg is for "grammar of graphics" - a declarative model of how you want the visualization to look.
Build up the vis in layers.

every plot in ggplot2 must have:
- data
- geometry
- aesthetics



## Layer 1 - data

ggplot (and everything else in the tidyverse) requires "tidy data".

Tidy means a data frame:

1. Each observation has exactly one row
2. Each feature is a column
3. Each value is in exactly one cell

Tidy data is "3rd normal form" in Codd's relational data theory.

```{r}
library(ggplot2)

earn19
```

An example that came up for a student in the last project:

```{r}
infections = read.csv("infections.csv", row.names = 1)

boxplot(infections$Circle.results)
```
Fixing this data format

1. Each observation has exactly one row
2. Each feature is a column
3. Each value is in exactly one cell

123  go
more rows or less rows?

WANT:

shape   | count
-----   | -----
circle  | 129
rectangle  | 84
line  | 234
normal  | 212
circle  | 240
...

```{r}
# need to pivot this table
inf2 = pivot_longer(infections, cols = everything(),
                    names_to = "shape",
                    values_to = "count")

ggplot(data = inf2, aes(x = shape, y = count)) + geom_boxplot()

# Won't work
#ggplot(data = infections, aes(x = shape, y = count)) + geom_boxplot()
```

tbl, data frame, table, are mostly interchangable

## Layer 2 - geometry

```{r}
ggplot(earn19) + geom_line()
```


## Layer 3 - aesthetics

123 go: for a line plot, is it OK for x to be repeated?

```{r}
ggplot(earn19, aes(x = quarter, y = median_weekly_earn, linetype = sex)) + geom_line()

# What if I wanted color to vary by sex, instead of linetype?
ggplot(earn19, aes(x = quarter, y = median_weekly_earn, color = sex)) + geom_line()
```

Suppose I want everything blue.

```{r}
ggplot(earn19, aes(x = quarter, y = median_weekly_earn, linetype = sex)) + geom_line(color = "blue")
```


## Layer 4 - scales

```{r}
ggplot(earn19, aes(x = quarter, y = median_weekly_earn, linetype = sex)) +
  geom_line(color = "blue") +
  labs(y = "median weekly earnings ($USD)", title = "Earnings by sex", 
       caption = "Data from Nick Ulle's R Basics")
```

## Annotations

```{r}
p = ggplot(earn19, aes(x = quarter, y = median_weekly_earn, linetype = sex)) +
  geom_line(color = "blue") +
  labs(y = "median weekly earnings ($USD)", title = "Earnings by sex", 
       caption = "Data from Nick Ulle's R Basics")

p = p + annotate("text", label = "elbow", x = 3, y = 1000, vjust = "top")
p
```



## Saving

```{r}
ggsave("line.png")
```

Plot devices

```{r}
```


