
How was the DS panel?

Which data sounds most interesting to you?

1. COVID cases
2. public financial records
3. news [GDELT](https://www.gdeltproject.org/)
4. census data


## ggplot2

Notes: [datalab](https://ucdavisdatalab.github.io/workshop_r_basics/exploring-your-data.html#data-visualization)

Authoritative reference: [ggplot2 book](https://ggplot2-book.org/introduction.html)

> "ggplot2 provides beautiful, hassle-free plots that take care of fiddly details like drawing legends. 
> ...
> rather than spending your time making your graph look pretty, you can instead focus on creating the graph that best reveals the message in your data."


Three popular systems for creating visualizations in R:

The base R functions
The lattice package
The ggplot2 package



```{r}

```



```{r}
earn = read.csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-23/earn.csv")
```

Let’s make a line plot that shows median earnings for each quarter in 2019, with separate lines for men and women.

```{r}

```

The tidyverse way:

```{r}

```




## Layer 1 - data


```{r}
library(ggplot2)

```

An example that came up for a student in the last project:

```{r}
infections = read.csv("infections.csv", row.names = 1)

```



## Layer 2 - geometry

```{r}
```


## Layer 3 - aesthetics

```{r}
```


```{r}
```

## Layer 4 - scales

```{r}

```

## Annotations

```{r}
```



## Saving

```{r}
ggsave("line.png")
```

Plot devices

```{r}
```


